---
title: "Vignette IndRSA"
author: "Guillaume Bastille-Rousseau"
date: "July 19, 2021"
output:
  pdf_document: default
  html_document: default
---


``` {r eval=T, echo=F, message=F, warning=F}
#library(devtools)
#install_github("BastilleRousseau/IndRSA")
library(IndRSA)
library(survival)
simu_coefs<-function(coef, se, n=1000) {
  out<-array(NA, c(nrow(coef), ncol(coef), n))
  for (i in 1:nrow(coef)) {
    for (j in 2:(ncol(coef)-2)) {
      out[i,j,]<-rnorm(n,coef[i,j], se[i,j])
    }}
  return(out)
}
simu_avg<-function(simu) {
 tt<-(matrix(unlist(lapply(1:dim(simu)[3], function(x) colMeans(simu[,,x], na.rm=T))), nrow=dim(simu)[3], ncol=dim(simu)[2], byrow=T))
}



simu_spe<-function(simu) {
  return(matrix(unlist(lapply(1:dim(simu)[3], function(x) apply(simu[,,x], 2, function(y) mean(abs(y), na.rm=T)))), nrow=dim(simu)[3], ncol=dim(simu)[2], byrow=T))
}

simu_sd<-function(simu) {
  return(matrix(unlist(lapply(1:dim(simu)[3], function(x) apply(simu[,,x], 2, function(y) sd(y, na.rm=T)))), nrow=dim(simu)[3], ncol=dim(simu)[2], byrow=T))
}


simu_cons2<-function(simu, col1, col2) {
  return(unlist(lapply(1:dim(simu)[3], function(x)   mean(abs(simu[,col1,x]-simu[,col2,x]), na.rm=T)  )))
}

simu_cons3<-function(simu, col1, col2, col3) {
  return(unlist(lapply(1:dim(simu)[3], function(x)   mean(abs(simu[,col1,x]-simu[,col2,x])+abs(simu[,col1,x]-simu[,col3,x])+abs(simu[,col3,x]-simu[,col2,x])/3, na.rm=T) )))
}

simu_rev2<-function(simu, col1, col2) {
  return(unlist(lapply(1:dim(simu)[3], function(x) mean(ifelse(sign(simu[,col1,x])!= sign(simu[,col2,x]), 1, 0), na.rm=T))))
}

simu_rev3<-function(simu, col1, col2, col3) {
  return(unlist(lapply(1:dim(simu)[3], function(x) mean(c(ifelse(sign(simu[,col1,x])!= sign(simu[,col2,x]), 1, 0), ifelse(sign(simu[,col1,x])== sign(simu[,col3,x]), 1, 0), ifelse(sign(simu[,col3,x])== sign(simu[,col2,x]), 1, 0)),na.rm=T))))
}

```

# Individual and population-level RSF - *rsf_ind* and *pop_avg*

The function *rsf_ind* performs individual-level RSFs. The function takes a list of candidate models (using *as.formula()*). The function *aictab_ind* performs population-level AIC model selection (by adding up individual model loglikelihood). *pop_avg* performs population-level averaging. Two methods are available for calculating confidence intervals, the default (and recommended) is based on Murtaugh (2007). A bootstrap approach (based on Prokopenko 2016) is also available. *ind_coef* and *ind_se* also extract individual-level coefficients and standard errors.<br> <br>
Murtaugh, P. Simplicity and complexity in ecological data anlysis. Ecology 88, 56–62 (2007).<br> <br>
Prokopenko, C. M., Boyce, M. S. & Avgar, T. Characterizing wildlife behavioural responses to roads using integrated step selection analysis. J. Appl. Ecol. 1, (2016).<br> <br>
\newline

``` {r eval=T}
data(goats)
ls1<-list()
ls1[[1]]<-as.formula(STATUS~ELEVATION+SLOPE+ET+ASPECT+HLI+TASP)
ls1[[2]]<-as.formula(STATUS~ET+ASPECT+HLI+TASP)
out<-rsf_ind(goats$ID, data=goats, form_ls=ls1)
aictab_ind(out) # Model 1 is the best 

#Calculate population average for the first model of the list (m=1)
pop<-pop_avg(m=1, out, method="murtaugh") 
pop[[1]] #Population average summary 
pop[[2]] #Individual level coefficients 
ind_coef(m=1, out) #Another way of getting individual coefficients 
ind_se(m=1, out) #Individual level standard errors 
kfold_ind(m=1, out, ls=ls1, grph=F) #kfold cross validation for each individual  
```

# Individual and population-level SSF  - *ssf_ind* and *pop_avg*
The function *ssf_ind* performs individual-level SSFs similar to *rsf_ind*. The function takes a list of candidate models (using *as.formula()*). The function *aictab_ind*, *pop_avg*, *ind_coef*, and *ind_se* works similarly than in example above. For the example below, we created a new column strata in the goat dataset that assigns two random points to each used points to create a conditional design similar to an SSF <br> <br>
\newline

``` {r eval=T}
data(goats)
goats<-goats[order(goats$ID),]
goats_use<-goats[goats$STATUS==1,]
goats_use$strata<-1:nrow(goats_use)
goats_rnd<-goats[goats$STATUS==0,]
goats_rnd$strata<-rep(1:nrow(goats_use), each=2)
goats_ssf<-rbind(goats_use, goats_rnd)
ls1<-list()
ls1[[1]]<-as.formula(STATUS~ELEVATION+SLOPE+ET+ASPECT+HLI+TASP+strata(strata))
ls1[[2]]<-as.formula(STATUS~ET+ASPECT+HLI+TASP+strata(strata))
out<-ssf_ind(goats_ssf$ID, data=goats_ssf, form_ls=ls1)
aictab_ind(out)
pop<-pop_avg(m=1, out, method="murtaugh")
pop[[1]] #Population average summary 
pop[[2]] #Individual level coefficients 
coef<-ind_coef(m=1, out)
se<-ind_se(m=1, out)

```



# Metrics of variation in resource selection behavior (specialization, heterogeneity, consistency and reversal)  - *simu_spe*, *simu_sd*, *simu_consX*, and *simu_revX*
The functions *simu_spe* and *simu_sd* calculates specialization and heterogeneity metrics as described in Bastille-Rousseau & Wittemyer (in review). To propagate uncertainty associated to individual coefficients, *simu_coefs* is first used to simulate individual coefficients based on the coefficient value and its standard error. <br> <br>
\newline

``` {r eval=T}
data(goats)
ls1<-list()
ls1[[1]]<-as.formula(STATUS~ELEVATION+SLOPE+ET+ASPECT+HLI+TASP)
out<-rsf_ind(goats$ID, data=goats, form_ls=ls1)
coef<-ind_coef(m=1, out)
se<-ind_se(m=1, out)
simu<-simu_coefs(coef, se, n=100)

#Specialization 
spe<-simu_spe(simu)
colnames(spe)<-names(coef)
head(spe)
apply(spe, 2, quantile, na.rm=T) #Show variation around estimate of each covariate
colMeans(spe) #Calculate average specialization for each covariate

#Heterogeneity
sd<-simu_sd(simu)
colnames(sd)<-names(coef)
head(sd)
apply(sd, 2, quantile, na.rm=T) #Show variation around estimate of each covariate
colMeans(sd) #Calculate average heterogeneity for each covariate

```

For consistency and reversal, calculations are done one covariate at a time. *simu_cons2* and *simu_rev2* are used for calculations when there are two time periods and *simu_cons3* and *simu_rev3* are used when there is three temporal periods. For the example below, we created an artificial Season column to the goat dataset to estimate temporal consistency and reversal.<br> <br>
\newline

``` {r eval=T}
data(goats)
goats$Season<-c("1", "2") #Adding a fake season column 
ls1<-list()
ls1[[1]]<-as.formula(STATUS~(ELEVATION+SLOPE+ET+ASPECT+HLI+TASP):Season)  
out<-rsf_ind(goats$ID, data=goats, form_ls=ls1)
coef<-ind_coef(m=1, out)
se<-ind_se(m=1, out)
simu<-simu_coefs(coef, se, n=100)

#Consistency for elevation 
head(coef)
#Calculate specialization for elevation covariate, column 3 and 4 contains 
#coefficients for elevation for each season 
cons_elevation<-simu_cons2(simu, 3, 4) 
quantile(cons_elevation) #Show variation around estimate of elevation covariate
mean(cons_elevation) #Calculate average consistency for elevation covariate

#Reversal for elevation 
head(coef)
rev_elevation<-simu_rev2(simu, 3, 4) #Calculate specialization for elevation covariate
quantile(rev_elevation) #Show variation around estimate of elevation covariate
mean(rev_elevation) #Calculate average reversal for elevation covariate

```


